---
title: "fit-gamma"
author: "ZeZhi Wang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fit-gamma}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(StatComp21077)
```

## Overview

__StatComp21077__ is a simple R package developed to fit gamma model which is an example of generalized linear model. We implement two methods for fitting, and a function to generate simulated data. Here, we will compare the performance and accuracy of them and __stats::glm__ function. Two functions are considered, namely, __fit.gamma.IWLS__ (fit the gamma model with IWLS method) and __fit.gamma.appro__ (fit the gamma model with approximate newton method). In addition, the needed simulated data will be generate by __generate.data__.

The R package 'microbenchmark' can be used to benchmark the above functions.

## Run code

It is worth mentioning that __stats::glm__ will report "Error: no valid set of coefficients has been found: please supply starting values" when fitting gamma model if the the start value isn't specified. However, there is no such problem in the both function __fit.gamma.IWLS__ and __fit.gamma.appro__ given by __StatComp21077__.


```{r,eval=TRUE}
library(microbenchmark)
dataset <- generate.data(1000, 10, 10,family = "gamma", seed = 1)

start_point <- c(abs(min(dataset[["x"]] %*% rep(1, 10))) + 1, rep(1, 10))
tm <- microbenchmark(
  coef.appro <- fit.gamma.appro(dataset[["x"]],dataset[["y"]]),
  coef.IWLS <- fit.gamma.IWLS(dataset[["x"]],dataset[["y"]]),
  coef.appro.start_point <- fit.gamma.appro(dataset[["x"]],dataset[["y"]],start = start_point),
  coef.IWLS.start_point <- fit.gamma.IWLS(dataset[["x"]],dataset[["y"]],start = start_point),
  coef.glm <- glm(y ~ ., data = cbind.data.frame("y" = dataset[["y"]],dataset[["x"]]), family = "Gamma", start = start_point)
)
```

## Performance

```{r,eval=TRUE}
name <- c("appro","IWLS","appro with start point","IWLS with start point","glm with start point")
knitr::kable(cbind(name,summary(tm)[,c(2,4,5)]))
```

We can easily find that the both methods are faster than __stats::glm__ whether the start value is specified or not.

## Accuracy

```{r,eval=TRUE}
mse <- function(est, true){
  mean((est - true)^2)
}
true <- coef(coef.glm)
deviation <- round(c(
  mse(coef.appro,true),
  mse(coef.IWLS,true),
  mse(coef.appro.start_point,true),
  mse(coef.IWLS.start_point,true)
),3)

knitr::kable(cbind(name[-5],deviation))
```


We can find that the results are similar with __stats::glm__.



